---
title: "Lab 09"
author: ""
output: html_notebook
---

The format of this lab follows the same format as the previous
ones. Your goal is to predict the value of the third column
(which will be missing on the test set) using the techniques
we have learned so far.

# Set up

Read in the following libraries and to load the crimes dataset:

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(viridis)

tract <- read_csv("https://statsmaths.github.io/ml_data/tract_median_income.csv")
```

This lab will be scored using RMSE.


```{r}
if (!require("glmnet")) {
  install.packages("glmnet")
}
```

# Lab 09

## Model selection and LM

Because it is the most recent thing we've learned, let's try to use the `cv.glmnet` function
to help with model selection. Here, it suggests that the the income quartiles are most predictive,
but also that housing costs are useful. This should probably not be surprising!

```{r}
library(glmnet)
X <- model.matrix(~ . -1, data = tract[,10:70])
y <- tract$median_income
X_train <- X[tract$train_id == "train",]
y_train <- y[tract$train_id == "train"]

model <- cv.glmnet(X_train, y_train, alpha = 0.7)
coef(model)
```

I suspect that I can get a very accurate model just using these five variables. Let's
try to do this with the data:

```{r}
model <- lm(median_income ~ income_q1 + income_q2 + income_q3 + income_q4,
            data = tract,
            subset = (train_id == "train"))
tract$median_income_pred_1 <- predict(model, newdata = tract)
sqrt(tapply((tract$median_income_pred_1 - tract$median_income)^2, tract$train_id, mean))
```

Compared to the mean estimator, this is really great great

```{r}
sqrt(tapply((mean(tract$median_income, na.rm = TRUE) - tract$median_income)^2, tract$train_id, mean))
```

Though, as we would expect, it is not too hard to predict tract median income if we have the
second and third income quartiles.

Perhaps we could get a better model by just using the second and third quartiles. With these,
there is also no danger of over-fitting:

```{r}
model <- lm(median_income ~ income_q2 + income_q3 - 1,
            data = tract,
            subset = (train_id == "train"))
tract$median_income_pred_2 <- predict(model, newdata = tract)
sqrt(tapply((tract$median_income_pred - tract$median_income)^2, tract$train_id, mean))
```

I am going to blend these two models together:

```{r}
tract$median_income_pred <- 0.9 * tract$median_income_pred_1 + 0.1 * tract$median_income_pred_2
sqrt(tapply((tract$median_income_pred - tract$median_income)^2, tract$train_id, mean))
```


# Submission

The code below assumes that you have added a prediction named
`median_income_pred` to every row of the dataset.

```{r}
submit <- select(tract, obs_id, median_income_pred)
write_csv(submit, "class09_submit.csv")
```

Now, upload this file (ends with ".Rmd"), the HTML output
(ends with ".nb.html" or ".html"), and the csv file to
GitHub.
